Union and intersection contracts are hard, actually

### 2021-06-09 paper.tex
Section 4 intro + random stuff

For instance, Figure~\ref{fig:port-contract} shows the definition of a
contract for valid ports in Nickel syntax.

@@

User-defined contracts can be combined with other contracts normally,
for instance \nickel{Int -> Port} is a contract verified by functions
which, given an integer returns a valid port.

This type of contracts are present on many different languages,
for instance, the Eiffel programming language\cite{meyer1987eiffel}, the precursor
of the Design by Contract ideals, allows to assert
these kinds of expression as pre- and post-conditions on
functions and as invariants on classes\cite{EiffelDesignByContract}.

@@

\section{Issues coming from the semantics}
\label{sec:issues-sem}
\info{What goes wrong or difficult when trying to implement union and
  intersection literally}

Static typing is about enforcing that some property about a (potentially
unknown) value always holds, no matter the execution path. A dynamic check, on
the other hand, just needs to verify a property for a specific and known value,
which is usually much easier. For example, dependent function types are complex
and mostly implemented in proof assistants (although they are slowly making
their way in some mainstream languages), whereas the corresponding dependent
function contracts can be implemented in a straightforward way (they are
supported in e.g. Racket~\cite{}).

The gist of the present paper is to bring evidence that in the case of unions
and intersections, the dynamic checking part is complex as well: \emph{unions
and intersections are fundamentally hard}.

% Teo's version
% Implementing a static type system is not an easy task.
% Type systems featuring intersection types are hard in a fundamental way,
% from their discovery, they've been studied as a form to characterize
% lambda-calculus terms normalizability; which means that typechecking a
% term on a system with intersection types is undecidable,
% similar to how the halting problem is undecidable\todo{reference}.
%
% Usually, when thinking about programming languages, there's an intuition
% that static checks (like typechecking) are more difficult and less powerful
% than dynamic checks (like testing).
% For instance, it is very easy to check if,
% in a given execution of a program, a variable holds an integer value; while it
% may become more challenging to check if a variable will always hold an integer value,
% no matter the particular execution.
%
% As we intend to show on this paper, unions and intersections are not only
% statically hard to check, but they are also dynamically hard:
% \emph{they are fundamentally hard}.
% In this section, we present intuitively the main challenges that arise
% when attempting to develop a runtime check system for unions and intersections.

>>>>

For instance, Figure~\ref{fig:port-contract} shows the definition of a
contract for valid ports in Nickel syntax.
%
User-defined contracts can be combined with other contracts normally:
\nickel{Int -> Port} is a contract verified by functions
which, given an integer returns a valid port.

@@

This type of contracts are present in many different languages,
for instance, the Eiffel programming language\cite{meyer1987eiffel}, the precursor
of the Design by Contract philosophy, makes it possible to assert
these kinds of expression as pre- and post-conditions on
functions and as invariants on classes\cite{EiffelDesignByContract}.

@@

\section{Incompatibilities}
\label{sec:issues-sem}

However appealing union and intersection contracts may be, they happen
to be fundamentally incompatible with the desirable language
features from Section~\ref{sec:feat-lang}. At least in there
full-blown form: in Section~\ref{sec:issues-literature} we will
discuss pragmatic restrictions of union and intersection contracts to
recover some or all of the features.
